---
title: "Coronavirus"
author: "G.Robertson"
date: "03/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coronavirus stats


Required libraries

```{r}
library(tidyverse)
library(rvest)
```

Scrape the data from worldometers.com
```{r}
url <- "https://www.worldometers.info/coronavirus/"

all <-url%>%read_html()%>%html_table()%>%.[[1]]
```

Clean up the data
```{r}
# remove the continents and "world"
coronavirus <- all[9:230,-1]

# Remove symbols from formatting of numbers
coronavirus[]<-lapply(coronavirus, function(x) (gsub("\\,|\\+", "", (x))))

# convert all but the first and last column to numeric
coronavirus[,c(2:14, 16:21)] <- sapply(coronavirus[c(2:14, 16:21)],as.numeric)

# assign appropriate column names
colnames(coronavirus) <- c("Country", "TotalCases", "NewCases", "TotalDeaths", "NewDeaths", "TotalRecovered", "NewRecovered", "ActiveCases", "CriticalCases", "Cases_Per1M", "Deaths_per1M", "TotalTests", "Tests_per1M", "Population", "Continent", "1CaseeveryXppl", "1DeatheveryXppl", "1TesteveryXppl", "NewCases_per1M", "NewDeaths_per1M", "ActiveCases_per1M")

```

Who has had it really bad?
```{r}
# deaths per million
a <- coronavirus[!is.na(coronavirus$Deaths_per1M),]

# Examine this variable to see what would be appropriate to extract. Don't want too many data points and dont want too few
range(a$Deaths_per1M)
# Extract the required range
topdeathrate <- a[a$Deaths_per1M>=2000, ]
```

```{r}
# Plot it
ggplot(topdeathrate, aes(x=reorder(Country, Deaths_per1M), y=Deaths_per1M, fill=Continent))+
      geom_bar(stat = "identity")+
      xlab("Country")+
      ylab("Deaths per 1,000,000")+
      theme(axis.text.x = element_text(angle = 90))+
      coord_flip()
```

Where are ases growing rapidly?
```{r}
b <- coronavirus[!is.na(coronavirus$NewCases_per1M),]

surge <- b[b$NewCases_per1M>=150,]    
```

Plot
```{r}
ggplot(surge, aes(x=reorder(Country, NewCases_per1M), y=NewCases_per1M, fill=Continent))+
      geom_bar(stat = "identity")+
      xlab("Country")+
      ylab("New cases per 1,000,000")+
      theme(axis.text.x = element_text(angle = 90))+
      coord_flip()
```

############################################

Islands are always kind of weird because of a variety of reasons:
Small populations make it easier to achieve 100% testing, so cases often appear higher compared to populous countries where testing % is much much lower. If you have a very high
rate of testing, you catch all the asymptomatic people, whereas low testing rates tend to prioritise symptomatic patients. Also,some asymptomatic testing in populous countries may rely on self-testing and self-reporting, which means some people simply won't bother.

So it might be a good idea to remove islands from the dataset
#############################################

Some random subsets
```{r}
continents <- all[1:6,]

# Low population countries, like islands etc.
lowpop <- coronavirus[coronavirus$Population<=400000,]
```

