---
title: "CoronaMapping"
author: "G.Robertson"
date: "01/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping current coronavirus information from Worldometers

Packages needed
```{r}
library(tidyverse)
library(rvest)
library(rgdal)
library(RColorBrewer)
```

Webscrape the data
```{r}
url <- "https://www.worldometers.info/coronavirus/"
all <-url%>%read_html()%>%html_table()%>%.[[1]]
```

Tidy it up
```{r}
# exclude the continents and world variables
coronavirus <- all[9:230,-1]

# Remove symbols from formatting of numbers
coronavirus[]<-lapply(coronavirus, function(x) (gsub("\\,|\\+", "", (x))))

# convert all but the first and last column to numeric
coronavirus[,c(2:14, 16:21)] <- sapply(coronavirus[c(2:14, 16:21)],as.numeric)

# rename the cols
colnames(coronavirus) <- c("Country", "TotalCases", "NewCases", "TotalDeaths",       
                           "NewDeaths", "TotalRecovered", "NewRecovered", "ActiveCases",       
                           "CriticalCases", "Cases_Per1M", "Deaths_per1M", "TotalTests",         
                           "Tests_per1M", "Population", "Continent", "1CaseeveryXppl", 
                           "1DeatheveryXppl",  "1TesteveryXppl",   "NewCases_per1M",
                           "NewDeaths_per1M", "ActiveCases_per1M")
```

Load a map file, plot it, and tweak the formatting
```{r}
map_object <- readOGR("C:/Users/georg/Documents/R_Maps/TM_WORLD_BORDERS_SIMPL-0.3.shp")
     
euro_map <- map_object[map_object@data$REGION==150, ]


plot(euro_map, xlim=c(30,50) , ylim=c(35, 75))
```

Data wrangling
```{r}
# subset europe data from main table
europe <- coronavirus[coronavirus$Continent=="Europe",]
# = 48 obs

# how many in the map data?
length(euro_map$NAME)
# = 51

# what are the discrepancies?
setdiff(euro_map$NAME, europe$Country)
# Listed below is what occurs in map data, but not in our corona data
```

```{r}
# Seeing how the "missing" countries are named in the corona data
europe$Country
```
# rename some of the variables in line with the map table convention
```{r}
europe$Country[europe$Country=="UK"] <- "United Kingdom"
europe$Country[europe$Country=="Czechia"] <- "Czech Republic"
europe$Country[europe$Country=="North Macedonia"] <- "The former Yugoslav Republic of Macedonia"
europe$Country[europe$Country=="Faeroe Islands"] <- "Faroe Islands"
europe$Country[europe$Country=="Moldova"] <- "Republic of Moldova"
europe$Country[europe$Country=="Vatican City"] <- "Holy See (Vatican City)"
```

```{r}
setdiff(euro_map$NAME, europe$Country)
# Just going to ignore these islands for now, sorry Islanders!
```


```{r}
# actually prefer this nomenclature
euro_map[euro_map$NAME=="Holy See (Vatican City)"] <- "Vatican City"
```

Merge the data
```{r}
euro_obj <- merge(euro_map, europe, by.x="NAME", by.y="Country")

head(euro_obj)
# success!
```
#### New cases per 1M

Plot a histogram of the data to determine an appropriate bin size
```{r}
euro_obj@data %>%
      ggplot(aes(x=as.numeric(NewCases_per1M))) +
      geom_histogram(bins = 15, fill="darkslategrey", color="white")
```
Get a colour palette sorted out
```{r}
par(mar=c(2,4,2,2)) # bottom, left, top, right
display.brewer.all()
```
```{r}
# define a palette
my_colours <- brewer.pal(7,"Oranges")
my_colours <- colorRampPalette(my_colours)(15)

# Attribute the appropriate color to each country
class_of_country <- cut(euro_obj@data$Cases_Per1M, 15)
my_colours <- my_colours[as.numeric(class_of_country)]

# Plot it
plot(euro_obj, xlim=c(40,55) , ylim=c(35, 75), col=my_colours,  bg = "#A6CAE0", main=paste("New cases per 1M as of ", Sys.Date()))


```

### Deaths per 1M

```{r}
# define a pallette
my_colours <- brewer.pal(7,"Oranges")
my_colours <- colorRampPalette(my_colours)(15)

# assign the colours to the countries
class_of_country <- cut(euro_obj@data$Deaths_per1M, 15)
my_colours <- my_colours[class_of_country]

# plot it
plot(euro_obj, xlim=c(40,55) , ylim=c(35, 75), col=my_colours,  bg = "#A6CAE0", main=paste("Deaths per 1M as of ", Sys.Date()))
```

